Automatic Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{Research supported by NSF OAC \#1940233 and \#1940322.} }

### 2021-04-05 paper/paper_new.tex
Updates from Overleaf

\section{Introduction}
% Moving teaser since this template doesn't actually ask for one
Over the last several decades advances in computing, imaging, and cyberinfrastructure have supported the growth of digital natural history collections, many of which have specimen images.~\cite{beaman2012mass} Additionally, initiatives, such as National Science Foundation’s Advancing Digitization of Biodiversity Collections (ADBC) program, which ran for nearly a decade, have supported the digitization and curation of tens of thousands of biological specimens.~\cite{page2015digitization} These digitized specimen are generally accessible through global, open access repositories supporting digital library services, such as browsing, search and retrieval, and preservation. These digitized renderings of these rich collections permit researchers, educators, students, and the general public to examine collections on a previously unattainable scale. Moreover, the digitized instantiations present a pathway for applying machine learning and making new scientific discoveries.

Unfortunately, potential scientific advances are hindered by image quality
problems and the lack of accurate and pertinent metadata
associated with the image collections.
Poor quality images (e.g. low contrast, inadequate lighting,
out-of-focus or cluttered visual arrangements) provide inadequate input
for automated
image analysis and machine learning algorithms and lead to inferior
computational results.
In order to perform quantitative morphometric analysis of the specimens,
the physical scale of the images (pixels/inch) is needed; thus
requiring the ability to identify and measure rulers in the images.
Many specimen collections do include Darwin Core metadata,~\cite{biodiv_info_standards} detailing
specimen taxon, geographic location, and several other descriptive aspects.
Additionally, some digitization efforts record technical metadata, detailing imaging specifications. While these types of metadata are helpful for a
human examining several images at a time, they are insufficient for researchers seeking to apply computational methods %machine learning? or okay here
to examine thousands of images
to determine if, for example, a specific fish grows to different lengths
in different habitats, or the impact of temporal factors the size of a dorsal fin.
\begin{figure*}
  \centering
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100314}
  \hspace{1mm}
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100682}
  \hspace{1mm}
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100020}
  \caption{Three example images from the Illinois Natural History Survey (INHS) collection.}
  \label{fig:INHS_examples}
\end{figure*}

@@

\section{Related Work}
\subsection{Metadata for Natural History Image Collection}
%question/comment - we need to keep in mind we are deal with digitized specimens, not simply digital, in other words, the artifacts themselves are digital version of physical, not like an x-ray, where the image is digital from the start.  
A number of different metadata standards have been applied to support the description and access of digitized scientific images. The Darwin Core (DwC)~\cite{biodiv_info_standards}, developed specifically to describe biological diversity data is among the most popular standard for such efforts. It is an extension of the Dublin Core's DCMI Metadata Terms \cite{dc_terms}. The Audubon Core\cite{audub_core} supporting the discoverability, dissemination, and use of data relating to biological organisms, including 3-D digitized specimens, is an DwC extension that has become more popular. All of these descriptive standards and extensions include metadata properties for taxon, geographic location, and other important specimen content, and have been developed primarily from the eye of a human curator. In other words, their application anticipates that a curator or data entry staff will manually generate metadata, drawing from acquisition logs or original specimen labels. This task is labor intensive, and the metadata associated with each digital rendering is generally sparse and more prone to human error, placing limitations on a researcher seeking to apply machine learning to the image and the metadata for scientific research.

This limitation is magnified when trying to assess the actual quality of the digitized specimen. Descriptive oriented standards support search and retrieval, and the biodiversity community has advocated for data fitness standards \cite{chapman2020developing}. This point is also emphasized by Wieczorek\cite{wieczorek2012darwin} in their report on the variety of DwC metadata extensions to meet growing community concerns and needs, including data quality and fitness. Even so, metadata describing image quality is severely limited and generally missing. This point is addressed in detail in \cite{leipzig2021biodiversity}, serves as the rationale for Tulane University's effort to manually capturing content for 22 metadata properties characterizing digitized specimen image quality. Their work is being conducted in connection with the lager BGNN initiative, and the current process is costly in terms of labor and time, prone to human error and inconsistencies, and underscores the need to
explore automatic metadata generation methods.
%Fish species dataset in Pakistan Shah 2019 \cite{Shah2019FishPakFS}
\subsection{Automatic Metadata Generation}
Automatic metadata generation advance covering both descriptive and technical metadata are relevant to the research presented in this paper. Automatic metadata generation of descriptive bibliographic data has been a research focus for close to 20 years \cite{liddy2002automatic,greenberg2004metadata,cardinaels2005automating,paynter2005developing}. Researchers have applied support vector machine (SVM) approaches \cite{han2003automatic}, and associated networks to address sparse and incomplete metadata \cite{rodriguez2009automatic}, and various successes are integrated into day-to-day workflows. Heidorn, et al,\cite{heidorn2008automatic} demonstrated the use optical character recognition (OCR) to extract specimen information from the original typed and often hand-annotated labels that are digitized along with herbarium collection holdings. The extracted information was  encoded in the DwC metadata associated with the specimen's digitized rendering. There's also been some success with extracting descriptive cartographic information from maps \cite{manso2004automatic}. While descriptive metadata covers taxon, geographic location, and other important aspects, and may even record the image format; extensive use of automatic processes are still limited. More significantly, descriptive metadata does not not sufficiently addressed data quality.

Technical metadata, such as camera settings and temporal information (date and time) are automatically generated during a digitization sequence, following standards such as Exchangeable image file format (Exif)\cite{Exif}, the camera's technical metadata that is automatically captured and inserted into digital image files at the time of digitization. Some of this metadata may be useful when selecting a machine learning sample. A researcher may desire images captured chiefly with a certain aperture setting. Even so, the majority of automatically captured technical metadata associated with digitized specimens is also insufficient for computational research, leaving researchers to rely on manually generated descriptive metadata, which itself is sparse and prone to human error. Fish image analysis research, as reviewed below, provides an opportunity to address current metadata shortcomings and needs specific to the selection of high-quality digitized specimen images for neural networks and the application of machine learning.

@@

for well over two decades \cite{Zion2000InvivoFS,Saberioon2017ApplicationOM}. It is an important

@@

vector machine (MSVM) \cite{Vapnik1999AOS}.
Li and Hong \cite{Li2014IdentificationOF} computed eleven shape and color

@@

Our work stands apart from previous work in that we are
addressing a different need. Lack of metadata is impeding scientific studies.
Our focus and motivation are different.
Effectively doing the computational pre-processing
that will enable further downstream analyses.

@@

Digitized specimen accessible in open access repositories provide a rich, extensive data source machine-learning and scientific discovery, however, these resources remain largely untapped due image quality and metadata limitations. We are addressing this need by combining machine-learning and image informatics technqiues. The overall goal of our

@@

\subsection{Detectron}
A prerequisite task to performing any advanced metadata property generation
is finding the specimens (and other relevant objects) within the collection
images. Object detection has been a broadly active field of study in recent
years \cite{zou2019object}, and has resulted in a number of well-tested, purpose-built architectures. We elected to use Facebook AI Research's \verb|detectron| tool \cite{wu2019detectron2}, and specifically its implementation of the Mask R-CNN architecture \cite{he2018mask}, for object detection in our project.
\verb|detectron| is built on \verb|pytorch|~\cite{NEURIPS2019_9015} and provides a relatively straightforward method for training on COCO~\cite{DBLP:journals/corr/LinMBHPRDZ14} format datasets. It is able to handle an arbitrary number of object classes, and can classify an arbitrary number of object within a given image. 
For our project, we use it to identify five object classes: fish, fish eyes, rulers, and the numbers 2 and 3 on rulers,
as shown in Figure \ref{fig:teaser}.

>>>>

\section{Introduction}
% Moving teaser since this template doesn't actually ask for one
Over the last several decades advances in computing, imaging, and cyberinfrastructure have supported the growth of digital natural history collections, many of which have specimen images~\cite{beaman2012mass}. Additionally, initiatives, such as National Science Foundation’s Advancing Digitization of Biodiversity Collections (ADBC) program, which ran for nearly a decade, have supported the digitization and curation of tens of thousands of biological specimens~\cite{page2015digitization}. These digitized specimen are generally accessible through global, open access repositories supporting digital library services, such as browsing, search and retrieval, and preservation. These digitized renderings of these rich collections permit researchers, educators, students, and the general public to examine collections on a previously unattainable scale. Moreover, the digitized instantiations present a pathway for applying machine learning and making new scientific discoveries.

Unfortunately, potential scientific advances are hindered by image quality
problems and the lack of accurate and pertinent metadata
associated with the image collections.
Poor quality images (e.g. low contrast, inadequate lighting,
out-of-focus or cluttered visual arrangements) provide inadequate input
for automated
image analysis and machine learning algorithms and lead to inferior
computational results.
In order to perform quantitative morphometric analysis of the specimens,
the physical scale of the images (pixels/inch) is needed; thus
requiring the ability to identify and measure rulers in the images.
Many specimen collections do include Darwin Core metadata~\cite{biodiv_info_standards}, detailing
specimen taxon, geographic location, and several other descriptive aspects.
Additionally, some digitization efforts record technical metadata, detailing imaging specifications. While these types of metadata are helpful for a
human examining several images at a time, they are insufficient for researchers seeking to apply computational methods %machine learning? or okay here
to examine thousands of images
to determine if, for example, a specific fish grows to different lengths
in different habitats, or the impact of temporal factors the size of a dorsal fin.
\begin{figure*}
  \centering
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100314}
  \hspace{1mm}
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100682}
  \hspace{1mm}
  \includegraphics[width=.31\linewidth]{images/INHS_FISH_100020}
  \caption{Three example images from the Illinois Natural History Survey (INHS) collection.}
  \label{fig:INHS_examples}
\end{figure*}

@@

\section{Related Work}
\subsection{Metadata for Natural History Image Collection}
%question/comment - we need to keep in mind we are deal with digitized specimens, not simply digital, in other words, the artifacts themselves are digital version of physical, not like an x-ray, where the image is digital from the start.  
A number of different metadata standards have been applied to support the description and access of digitized scientific images. The Darwin Core (DwC)~\cite{biodiv_info_standards}, developed specifically to describe biological diversity data is among the most popular standard for such efforts. It is an extension of the Dublin Core's DCMI Metadata Terms~\cite{dc_terms}. The Audubon Core~\cite{audub_core} supporting the discoverability, dissemination, and use of data relating to biological organisms, including 3--D digitized specimens, is an DwC extension that has become more popular. All of these descriptive standards and extensions include metadata properties for taxon, geographic location, and other important specimen content, and have been developed primarily from the lens of a human curator. In other words, their application anticipates that a curator or data entry staff will manually generate metadata, drawing from acquisition logs or original specimen labels. This task is labor intensive, and the metadata associated with each digital rendering is generally sparse and more prone to human error, placing limitations on a researcher seeking to apply machine learning to the image and the metadata for scientific research.

This limitation is magnified when trying to assess the actual quality of the digitized specimen. Descriptive oriented standards support search and retrieval, and the biodiversity community has advocated for data fitness standards \cite{chapman2020developing}. This point is also emphasized by Wieczorek~\cite{wieczorek2012darwin} in their report on the variety of DwC metadata extensions to meet growing community concerns and needs, including data quality and fitness. Even so, metadata describing image quality is severely limited and generally missing. This point is addressed in detail in \cite{leipzig2021biodiversity}, serves as the rationale for Tulane University's effort to manually capturing content for 22 metadata properties characterizing digitized specimen image quality. Their work is being conducted in connection with the lager BGNN initiative, and the current process is costly in terms of labor and time, prone to human error and inconsistencies, and underscores the need to
explore automatic metadata generation methods.
%Fish species dataset in Pakistan Shah 2019 \cite{Shah2019FishPakFS}
\subsection{Automatic Metadata Generation}
Automatic metadata generation advances covering both descriptive and technical metadata are relevant to the research presented in this paper. Automatic metadata generation of descriptive bibliographic data has been a research focus for close to 20 years \cite{liddy2002automatic,greenberg2004metadata,cardinaels2005automating,paynter2005developing}. Researchers have applied support vector machine (SVM) approaches~\cite{han2003automatic}, and associated networks to address sparse and incomplete metadata~\cite{rodriguez2009automatic}, and various successes are integrated into day-to-day workflows. Heidorn, et al.~\cite{heidorn2008automatic} demonstrated the use of optical character recognition (OCR) to extract specimen information from the original typed and often hand-annotated labels that are digitized along with herbarium collection holdings. The extracted information was encoded in the DwC metadata associated with the specimen's digitized rendering. There has also been some success with extracting descriptive cartographic information from maps~\cite{manso2004automatic}. While descriptive metadata covers taxon, geographic location, and other important aspects, and may even record the image format; extensive use of automatic processes are still limited. More significantly, descriptive metadata does not not sufficiently addressed data quality.

Technical metadata, such as camera settings and temporal information (date and time) are automatically generated during a digitization sequence, following standards such as Exchangeable image file format (Exif)~\cite{Exif}. The camera's technical metadata that is automatically captured and inserted into digital image files at the time of capture. Some of this metadata may be useful when selecting a machine learning sample. A researcher may desire images with specific properties, such as being captured chiefly with a certain aperture setting. Even so, the majority of automatically captured technical metadata associated with digitized specimens is also insufficient for computational research, leaving researchers to rely on manually generated descriptive metadata, which itself is sparse and prone to human error. Fish image analysis research, as reviewed below, provides an opportunity to address current metadata shortcomings and needs specific to the selection of high-quality digitized specimen images for neural networks and the application of machine learning.

@@

for well over two decades~\cite{Zion2000InvivoFS,Saberioon2017ApplicationOM}. It is an important

@@

vector machine (MSVM)~\cite{Vapnik1999AOS}.
Li and Hong~\cite{Li2014IdentificationOF} computed eleven shape and color

@@

Our work stands apart from previous work in that we are
addressing a different need: lack of metadata is impeding scientific studies.
Our focus and motivation are different.
Effectively doing the computational pre-processing
that will enable further downstream analyses.

@@

Digitized specimen accessible in open access repositories provide a rich, extensive data source machine-learning and scientific discovery, however, these resources remain largely untapped due image quality and metadata limitations. The overall goal is addressing this need by combining machine-learning and image informatics techniques. The overall goal of our

@@

\subsection{Detectron}
A prerequisite task to performing any advanced metadata property generation
is finding the specimens (and other relevant objects) within the collection
images. Object detection has been a broadly active field of study in recent
years \cite{zou2019object}, and has resulted in a number of well-tested, purpose-built architectures. We elected to use Facebook AI Research's (FAIR) \verb|detectron| tool \cite{wu2019detectron2}, and specifically its implementation of the Mask R-CNN architecture \cite{he2018mask}, for object detection in our project.
\verb|detectron| is built on \verb|pytorch|~\cite{NEURIPS2019_9015} and provides a relatively straightforward method for training on COCO~\cite{DBLP:journals/corr/LinMBHPRDZ14} format datasets. It is able to handle an arbitrary number of object classes, and can classify an arbitrary number of object within a given image. We chose \verb|detectron| for its relative ease of use compared to lower level libraries, and its implementations of powerful architectures developed by FAIR.
For our project, we use it to identify five object classes: fish, fish eyes, rulers, and the numbers 2 and 3 on rulers,
as shown in Figure \ref{fig:teaser}.
